---
title: "datamations steps and conventions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{datamations steps and conventions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

datamations constructs one _or more_ frames for each step of a pipeline. For example, in the following pipeline:

```{r mean-salary-degree}
library(datamations)
library(dplyr)

"small_salary %>% 
  group_by(Degree) %>%
  summarize(mean = mean(Salary))" %>%
  datamation_sanddance()
```

there are three steps:

1. The initial data (`small_salary`)

    An information grid is shown, laying out the number of points in the data set.
    
2. The grouped data (grouped by `Degree`)

    The data is separated into groups, retaining the informaton grid structure.

3. The summarized data (mean of `Salary`)

    The distribution of `Salary` within the groups is shown, then the summary function (mean) is applied. Error bars are added to the mean and the final frame zooms in on the data.
    
## `group_by()` frames

datamations supports *up to three* grouping variables, showing one frame
per variable. The placement of the variables is as follows:

- **One variable**: On the x-axis
- **Two variables**: The first variable in column facets, the second on the x-axis
- **Three variables**: The first variable in column facets, the second in row facets, the third in on the x-axis    
    
## `summarize()` frames

datamations supports summarizing _one_ variable. The `summarize()` section of a pipeline will have the following frames:

1. Distribution of the variable to be summarized
2. Summarized variable
3. Summarized variable with standard error (only if summary function is mean)
4. Zoomed version of summarized variable

## `count()` frames

datamations treats `count()` equivalently to `group_by()` + `summarize(n = n())`. It supports up to three "grouping" variables.

## `filter()` frames

datamation supports `filter()` at any point in the pipeline, whether it comes after the initial data, while the data is grouped, or after it has been summarized. 
