---
title: "Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

```{r setup}
library(datamations)
library(dplyr)
```

# group_by()

## One grouping variable

```{r}
"small_salary %>%
  group_by(Work)" %>%
  datamation_sanddance()
```

## Two grouping variables

```{r}
"small_salary %>%
  group_by(Work, Degree)" %>%
  datamation_sanddance()
```

## Three grouping variables

```{r}
library(palmerpenguins)

"penguins %>%
  group_by(sex, island, species)" %>%
  datamation_sanddance()
```

# summarise()

## mean of a variable

```{r}
"small_salary %>%
  group_by(Work) %>%
  summarise(mean_salary = mean(Salary))" %>%
  datamation_sanddance()
```

## Other summary functions

```{r}
"small_salary %>%
  group_by(Work) %>%
  summarise(median_salary = median(Salary))" %>%
  datamation_sanddance()
```

```{r}
"small_salary %>%
  group_by(Degree) %>%
  summarise(quan = quantile(Salary, probs = 0.01))" %>%
  datamation_sanddance()
```

# filter()

## Filtering initial data

```{r}
"small_salary %>%
  filter(Salary > 90)" %>%
  datamation_sanddance()
```

## Filtering within groups

```{r}
"small_salary %>%
  group_by(Work) %>%
  filter(Salary == mean(Salary))" %>%
  datamation_sanddance()
```

## Filtering after summarize

```{r}
"small_salary %>%
  group_by(Work) %>%
  summarise(median_salary = median(Salary)) %>%
  filter(median_salary > 90)"  %>%
  datamation_sanddance()
```

# count()

```{r}
"small_salary %>%
  count(Work)" %>%
  datamation_sanddance()
```

# Binary variables

A basic example of a dataframe containing data that represents Simpson's paradox.

```{r} 
head(jeter_justice)
```


In this datamation, Jeter has a higher batting average than Justice overall.

```{r}
'jeter_justice %>%
  group_by(player) %>%
  summarize(batting_average = mean(is_hit),
            se = sqrt(batting_average * (1 - batting_average) / n()) )' %>%
  datamation_sanddance()
```
<br>
<br>
In this datamation, the visual shows that Justice has a higher batting average than Jeter within each year.

```{r}
'jeter_justice %>%
  group_by(player, year) %>%
  summarize(batting_average = mean(is_hit),
            se = sqrt(batting_average * (1 - batting_average) / n()) )' %>%
  datamation_sanddance()
```
