---
title: "Controlling appearance with ggplot2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Controlling appearance with ggplot2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)

ggplot2::theme_set(ggplot2::theme_minimal())
```

If you would like to change the default conventions, or to match an existing plot style, datamations can take ggplot2 code as input.

For example, to match this plot, which has Work on the x-axis and Degree in row facets:

```{r ggplot2-existing-plot, dpi = 300}
library(datamations)
library(dplyr)
library(ggplot2)

small_salary %>%
  group_by(Work, Degree) %>%
  summarize(mean_salary = mean(Salary)) %>%
  ggplot(aes(x = Work, y = mean_salary)) + 
  geom_point() + 
  facet_grid(rows = vars(Degree))
```

Simply define the code and pass to `datamation_sanddance()`, which will produce an animation with desired plot layout.

```{r mean-salary-degree-work-ggplot-setup, eval = TRUE}
"small_salary %>%
  group_by(Work, Degree) %>%
  summarize(mean_salary = mean(Salary)) %>%
  ggplot(aes(x = Work, y = mean_salary)) + 
  geom_point() + 
  facet_grid(rows = vars(Degree))" %>%
  datamation_sanddance()
```

When ggplot2 code is provided, the order of animation is not determined by the order in `group_by()`, but by the plot layout. Variables are first animated by what's in the column facets, then the row facets, by the x-axis, and finally by color.

Some limitations:

* `facet_wrap()` is not supported - please use `facet_grid()`
* datamations expects different variables in the column and row facets than in the x-axis. datamations generated that do not match this layout may look different than the final plot!
* Only `geom_point()` is supported, e.g. specifying `geom_bar()` will not produce a bar in the datamation.
