---
title: "datamations group_by api"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{datamations group_by api}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

datamations allows the visualization of `dplyr::group_by` calls containing up to three variables.

## group_by api

The `datamations_sanddance` call can finish with a `group_by` call, displaying proportional data faceted by the grouping variables.

## Single grouping variable

```{r}
library(datamations)
library(dplyr)

"small_salary %>%
  group_by(Work, Degree)"
```
## Two grouping variables

```{r}
"small_salary %>%
  group_by(Work, Degree)" %>%
  datamation_sanddance()
```

## Three grouping variables

```{r}
library(palmerpenguins)

"penguins %>%
  group_by(sex, island, species)" %>%
  datamation_sanddance()
```

## summarizing and filtering by group

The `datamations_sanddance` call can also show the filtering and summarization of data based on group facets.

### group_by into filter

```{r}
"small_salary %>%
  group_by(Degree) %>%
  filter(Salary > 90)" %>%
  datamation_sanddance()
```

### group_by into summarize

```{r}
"small_salary %>%
  group_by(Degree) %>%
  summarize(mean = mean(Salary))" %>%
  datamation_sanddance()

"penguins %>%
  group_by(sex, island, species) %>%
  summarize(mean = median(bill_length_mm))" %>%
  datamation_sanddance()
```


