---
title: "datamations mutation api"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{datamations mutation api}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

{datamations} supports the limited definition of mutations in a pipeline. It is capable of showing a single mutation involving multiple variables in a scatterplot or grid fashion.

We can define new data to use in some example mutations.

### New data with added variable

```{r newdata}
library(dplyr)
library(datamations)

small_salary <- dplyr::mutate(
  small_salary, 
  supplementalIncome = runif(nrow(small_salary), min = 60, max = 110),
  logNorm = rlnorm(nrow(small_salary), meanlog = 0, sdlog = 1)
  )

```

{datamations} can visualize mutations to help one understand mathematical distributions, scales, and relationships. 

### Log normal mutation

```{r lognorm}

"small_salary %>%
  mutate(logged = log10(logNorm)) %>%
  group_by(Degree) %>%
  summarize(mean = mean(logged))" %>%
  datamation_sanddance()
```


### Mathematical Notation

```{r mutations}

"small_salary %>%
  mutate(salarySquared = Salary^2) %>%
  group_by(Degree) %>%
  summarize(mean = mean(salarySquared))" %>%
  datamation_sanddance()

"small_salary %>%
  mutate(inverseSalary = 1 / Salary) %>%
  group_by(Degree) %>%
  summarize(mean = mean(inverseSalary))" %>%
  datamation_sanddance()

```


### Multivariate mutates

{datamations} can also showcase the relationship between more than variable in your data pipelines. We can see below the relationship between our Salary variable and a new variable and use the mutation in grouping, filtering, and summarization.

```{r mutivar}

"small_salary %>%
  mutate(totalIncome = Salary + supplementalIncome) %>%
  group_by(Degree) %>%
  summarize(mean = mean(totalIncome))" %>%
  datamation_sanddance()

"small_salary %>%
  mutate(incomePer = Salary / supplementalIncome) %>%
  group_by(Degree) %>%
  summarize(mean = mean(incomePer))" %>%
  datamation_sanddance()

```

### Two variable mutates

{datamations} will allow the definition of a mutate statement with multiple mutates, but it will ignore anything after the first defined mutate. Two variable mutates results in a warning.

```{r twovar, error = TRUE}
"small_salary %>%
  mutate(totalIncome = Salary + supplementalIncome, squaredIncome = Salary^2) %>%
  group_by(Degree) %>%
  summarize(mean = mean(totalIncome))" %>%
  datamation_sanddance()
```
