---
title: "summarize, count, and tally"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{datamations steps and conventions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

The datamations API can visualize the summarization of variables with several dplyr verbs, like `summarize()`/`summarise()`, `count()`, and `tally()`

## summary functions

`datamations_sanddance` has special animations to visualize certain common summary functions passed to `summarize`. These custom animations include:

-   `mean`
-   `median`
-   `max`/`min`
-   `quantile`


### mean of a variable

```{r setup}
library(datamations)
library(dplyr)
```

```{r}
"small_salary %>%
  group_by(Work) %>%
  summarise(mean_salary = mean(Salary))" %>%
  datamation_sanddance()
```

### median of a variable

```{r}
"small_salary %>%
  group_by(Work) %>%
  summarise(median_salary = median(Salary))" %>%
  datamation_sanddance()
```

### min and max of a variable

```{r}
"small_salary %>%
  group_by(Work) %>%
  summarise(min_salary = min(Salary))" %>%
  datamation_sanddance()
```

```{r}
"small_salary %>%
  group_by(Work) %>%
  summarise(max_salary = max(Salary))" %>%
  datamation_sanddance()
```

### quantile() function

The summarize function includes the capacity to pass some parameterized functions that result in custom animations. Currently datamations supports the `quantile()` function with the `probs` parameter as an example of this capability. 

```{r}
"small_salary %>%
  group_by(Degree) %>%
  summarise(quan = quantile(Salary, probs = 0.01))" %>%
  datamation_sanddance()
```

## Other summary functions

Common summary functions may include count-like operations, like n_distinct.

### n_distinct()

```{r}
library(palmerpenguins)

"penguins %>%
  group_by(island) %>%
  summarise(n = n_distinct(species))" %>%
  datamation_sanddance()
```

You can find the API can support a large variety of mathematical functions and dplyr functions


```{r}
"small_salary %>%
  group_by(Degree) %>%
  summarize(floor_salary = trunc(Salary))" %>%
  datamation_sanddance()
```

```{r}
"small_salary %>%
  group_by(Degree) %>%
  summarize(first_salary = first(Salary))" %>%
  datamation_sanddance()
```

```{r}
"small_salary %>%
  group_by(Degree) %>%
  summarize(lagged_salary = lag(Salary))" %>%
  datamation_sanddance()
```

## count() and tally()

datamations treats `count()` and `tally()` calls equivalently to `group_by()` + `summarize(n = n())`.

```{r}
"small_salary %>%
  group_by(Degree) %>%
  count(Salary)" %>%
  datamation_sanddance()
```