---
title: "Variable types"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Variable types}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE
)
```

{datamations} provides useful visual encodings to help visualize transformations that occur in various types of data:

-   numeric
-   categorical
-   binary

## Numeric data

{datamations} can accept calculations on numeric data and provides encodings on faceted numeric scales. 

Test out the functionality on the built in dataset `small_salary`, which includes synthetic salary data for different types of work and different degree qualifications.

```{r setup}
library(datamations)
library(dplyr)
```

```{r}
"small_salary %>%
  group_by(Degree) %>%
  summarize(mean = mean(Salary, trim=0.1))" %>%
  datamation_sanddance()
```

## Categorical data

{datamations} generates shape encoding for categorical variables passed to `group_by` or in summary functions. A useful example is in the {palmerpenguins} `penguins` dataset.

```{r}
library(palmerpenguins)

"penguins %>%
  group_by(year, island) %>%
  summarise(n = n_distinct(species))" %>%
    datamation_sanddance()
```

## Binary data

{datamations} similarly can produce visualizations for binary outcomes, displaying variables encoded with opacity and stroke. Binary variables can be passed as a true binary class in R, e.g. `TRUE`/`FALSE`, or via a numeric variable encoded to `1`/`0`. This is examplified in the built in dataset `jeter_justice`, where outcome values are encoded as `1` and `0`.

```{r}
head(jeter_justice)
```

```{r}
"jeter_justice %>%
  group_by(year) %>%
  summarise(mean = mean(is_hit))" %>% datamation_sanddance()
```

The appearance of this encoding can also be controlled more finely with [ggplot aesthetics](https://microsoft.github.io/datamations/articles/finer_control.html).